Scoring Patients

### 2017-01-17 Clustering.Rmd
AP clustering analysis, as far as I’ve gotten so far

```{r setup, include=FALSE}
library(ggplot2)
library(GGally)
library(plyr)
library(xlsx)
library(dplyr)
library(pander)
library(Hmisc)
library(ggdendro)
library(reshape2)
library(GenomicRanges)

@@

## Segmentation Values
```{r clust, echo=T, message=T, warning=F}

@@

  model = Mclust(patient.data[[pt]]$seg.vals[,-(1:5)], modelNames=c('VEE','VVE'))
  print(model$loglik)

@@

## Normalized Segmentation Values
Doesn't work for the normalized values.  I'm not sure that the normalization is useful though...
```{r clustN, echo=T, message=T, warning=F}

  model = Mclust(patient.data[[pt]]$norm.seg.vals, modelNames=c('VEE','VVE'))
  print(paste('loglik', round(model$loglik,2)))

## Perhaps AP Clustering
```{r apclust-func, echo=F, message=F, warning=F}
suppressPackageStartupMessages( library(apcluster) )

apclust.data<-function(segdata, samples) {
  x1 = segdata[,intersect(colnames(segdata), samples)]
  rownames(x1) = (segdata[,c(1:4)] %>%
    rowwise() %>%
    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
  return(x1)  
}

@@

Affinity Propagagion clustering described by Frey & Dueck, 2007

```{r apclust, echo=T, message=T, warning=F, fig.height=5.5, fig.align='left'}
sum.patient.data = arrange(sum.patient.data, Status)

## Using the segmentation values
sim = sum.patient.data[,c(2:4)]
rownames(sim) = sum.patient.data[,1]
for (pt in sum.patient.data$Patient) {
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)

  sim[pt, 'ns'] = net.similarity(ac)
  sim[pt, 'clusters'] = length(clusters(ac))

  #plot(ac, x1, main=pt)
  #heatmap(ac, main=pt)
  heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'), main=pt)
}

summarise(group_by(sim, Status), mean.ns=mean(ns), sd.ns=sd(ns), mean.c=mean(clusters), sd.c=sd(clusters) )

# May be a difference
t.test( subset(sim, Status == 'NP')$ns, subset(sim, Status == 'P')$ns )
t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )

# Should this be performed on the normalized seg values instead?
sim3 = sum.patient.data[,c(2:4)]
rownames(sim3) = sum.patient.data[,1]
for (pt in sum.patient.data$Patient) {
  x1 = patient.data[[pt]]$norm.seg.vals
  ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)

  sim3[pt, 'ns'] = net.similarity(ac)
  sim3[pt, 'clusters'] = length(clusters(ac))

  #plot(ac, x1, main=paste(pt, 'norm'))
  heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'), main=paste(pt, 'norm'))
}

@@

# Doesn't show up here
t.test( subset(sim3, Status == 'NP')$ns, subset(sim3, Status == 'P')$ns )
t.test( subset(sim3, Status == 'NP')$clusters, subset(sim3, Status == 'P')$clusters )

@@

So using the segmentation values (not the normalized ones) try some leave one out xval to see if the currently weak differences still hold.

```{r xval, echo=T, warning=F, message=T, fig.align='left', fig.height=4}

# Leave out patients
loo_xval = matrix(ncol=2, nrow=0, dimnames=list(c(), c('net.similarity', 'num.clusters')))
for (i in 1:nrow(sum.patient.data)) {
  simL = sum.patient.data[-i,c(2:4)]
  rownames(simL) = sum.patient.data[-i,1]

  for (pt in sum.patient.data$Patient[-i]) {
    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
    ac = apcluster(negDistMat(x1, r=2), q=0.1)

    simL[pt, 'ns'] = net.similarity(ac)
    simL[pt, 'clusters'] = length(clusters(ac))
  }

  ttp = t.test( subset(simL, Status == 'NP')$ns, subset(simL, Status == 'P')$ns )$p.value
  ttc = t.test( subset(simL, Status == 'NP')$clusters, subset(simL, Status == 'P')$clusters )$p.value

  loo_xval = rbind(loo_xval, cbind(ttp,ttc))
}

ggplot(melt(data.frame(loo_xval)), aes(x=variable, y=value, fill=variable)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (patient) out clustering')

# Leaving out these patients causes the p-value to be > 0.05
pander(sum.patient.data[ which(loo_xval >= 0.05), ], caption='In a leave-one out analysis for ap clustering removing these patients causes the p-values to be > 0.05')

# They have a higher mean than average, but may not be significant
t.test(sum.patient.data[ which(loo_xval >= 0.05), 'n.samples'], subset(sum.patient.data, Status == 'P')$'n.samples')

# Leave out samples?
looS_xval = matrix(ncol=2, nrow=0, dimnames=list(c(), c('net.similarity', 'num.clusters')))
for (i in 1:20) { # 10 fold?
  print(i)
  simX = sum.patient.data[,c(2:4)]
  rownames(simX) = sum.patient.data[,1]
  # Leave one out for every clustering pass, but not first or last?
  for (pt in sum.patient.data$Patient) {
      x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-sample(2:(length(patient.data[[pt]]$info$Samplename)-1), 1)])
      ac = apcluster(negDistMat(x1, r=2), q=0.1)

      simX[pt, 'ns'] = net.similarity(ac)
      simX[pt, 'clusters'] = length(clusters(ac))
    }

  ttp = t.test( subset(simX, Status == 'NP')$ns, subset(simX, Status == 'P')$ns )$p.value
  ttc = t.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value

  looS_xval = rbind(looS_xval, cbind(ttp,ttc))
}

ggplot(melt(data.frame(looS_xval)), aes(x=variable, y=value, fill=variable)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (sample) out clustering')

## Shared gains/losses

Use ap clustering to identify the shared gains/losses in a single sample

```{r gl, echo=T, message=T, warning=F, fig.align='left', fig.height=10}

pt = 'PR1_ADH_074'

head(patient.data[[pt]]$variable.regions)

x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)

# hold on...perhaps I should compute the 
ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)

heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'))
#plot(ac, x1)
#plot(ac)

# Merge clusters
agex = aggExCluster(x=ac)
plot(agex)

plot(agex, x1, k=3)
heatmap(agex, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'))

ac
```

>>>>

```{r setup, include=FALSE}
library(ggplot2)
library(GGally)
library(plyr)
library(xlsx)
library(dplyr)
library(pander)
library(Hmisc)
library(ggdendro)
library(reshape2)
library(GenomicRanges)
library(gridExtra)
library(biomaRt)

@@

## How do the samples separate?

```{r pca, echo=F, warning=F, message=F, fig.width=20, fig.height=20}
sum.patient.data = arrange(sum.patient.data, Status, n.samples)

plot.pca<-function(pt) {

  info = arrange(patient.data[[pt]]$info, Endoscopy.Year)

  x1 = patient.data[[pt]]$seg.vals[, patient.data[[pt]]$info$Samplename]
  x1 = x1[,patient.data[[pt]]$info$Samplename]
  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )

  pca = prcomp(t(x1), scale=T)

  eigenvalues = pca$sdev^2
  #screeplot(pca, main='scree')

  #plot(pca$sd^2)

  df = as.data.frame(pca$x)
  df$year = as.factor(info$Endoscopy.Year)
  df$label = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )
  df$Pathology = info$Pathology

  gg = ggplot(df, aes(PC1, PC2)) + geom_line() +
    geom_label(aes(label=year),nudge_y=0.04) + 
    geom_point( aes(color=Pathology, size=Pathology)) +
    labs(title=pt)

  return(gg)
}

pplots = lapply(subset(sum.patient.data, Status == 'P')$Patient, plot.pca)
grid.arrange(grobs=pplots, ncol=2, top="Progressors")

npplots = lapply(subset(sum.patient.data, Status == 'NP')$Patient, plot.pca)
grid.arrange(grobs=npplots, ncol=2, top="Non-Progressors")

```{r clust, echo=T, message=T, warning=F, eval=F}

@@

  x1 = patient.data[[pt]]$seg.vals[,-(1:5)]

  model = Mclust(t(x1))
  print(model$loglik)

  tryCatch({
    mod1dr = MclustDR(model, lambda = 0)
    #summary(mod1dr)
    table(mod1dr$mixcomp)

@@

```
```{r apclust-func, echo=F, message=F, warning=F}
suppressPackageStartupMessages( library(apcluster) )

apclust.data<-function(segdata, samples) {
  x1 = segdata[,intersect(colnames(segdata), samples)]
  x1 = x1[, samples]
  rownames(x1) = (segdata[,c(1:4)] %>%
    rowwise() %>%
    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
  return(x1)  
}

@@

```{r em, echo=F, message=F, warning=F, eval=F} 
## EM

```{r clustN, echo=T, message=T, warning=F, eval=F}
## Normalized Segmentation Values
#Doesn't work for the normalized values.

  x1 = patient.data[[pt]]$norm.seg.vals
  model = Mclust(t(x1))
  print(paste('loglik', round(model$loglik,2)))

## Perhaps AP Clustering

## Affinity Propagagion clustering described by Frey & Dueck, 2007

First on the raw segmentation values

```{r apclust, echo=T, message=T, warning=F, fig.height=8, fig.align='left'}
q=0.2

sum.patient.data = arrange(sum.patient.data, Status)

## Using the segmentation values
sim = sum.patient.data[,c(2:4)]
rownames(sim) = sum.patient.data[,1]
for (pt in sum.patient.data$Patient) {
  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))

  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )

  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)

  last = ncol(x1)
  if (patient.data[[pt]]$info$Status == 'P')
    last = sort(grep('HGD', patient.data[[pt]]$info$Pathology), decreasing=T)[1]
  if (ncol(x1) <= 15) {
#    plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
  } else {
#    plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
  }

  sim[pt, 'ns'] = net.similarity(ac)
  sim[pt, 'ss'] = sum.similarity(ac)
  sim[pt, 'clusters'] = length(clusters(ac))
  #heatmap(ac)
}
sim$id = rownames(sim)

sm = as.data.frame(summarise(group_by(sim, Status), mean.ns=mean(ns), sd.ns=sd(ns), mean.ss=mean(ss), sd.ss=sd(ss), mean.c=mean(clusters), sd.c=sd(clusters) ))

# May be a difference
ttps = t.test( subset(sim, Status == 'NP')$ss, subset(sim, Status == 'P')$ss )$p.value
ttpn = t.test( subset(sim, Status == 'NP')$ns, subset(sim, Status == 'P')$ns )$p.value
ttpc = t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value

#t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value

g1 = ggplot(sim[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + 
  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
  labs(title=paste("p-value", signif(ttps, 3)), y='sum similarity')

g2 = ggplot(sim[,c('Status','clusters')], aes(y=clusters, x=Status, fill=Status)) + geom_boxplot() + 
  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
  labs(title=paste("p-value", signif(ttpc, 3)), y='num. clusters')

grid.arrange(g1, g2, top='Raw seg values')

So what about using the normalized segmented values?

```{r apclustN, echo=T, message=T, warning=F, fig.height=8, fig.align='left'}
# Should this be performed on the normalized seg values instead?
sim3 = sum.patient.data[,c(2:4)]
rownames(sim3) = sum.patient.data[,1]
for (pt in sum.patient.data$Patient) {
  x1 = patient.data[[pt]]$norm.seg.vals
  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)
  #plot(ac)

  sim3[pt, 'ss'] = sum.similarity(ac)
  sim3[pt, 'ns'] = net.similarity(ac)
  sim3[pt, 'clusters'] = length(clusters(ac))
}

@@

# Doesn't show up here
ttp = t.test( subset(sim3, Status == 'NP')$ss, subset(sim3, Status == 'P')$ss )$p.value
#ttp = t.test( subset(sim3, Status == 'NP')$ns, subset(sim3, Status == 'P')$ns )$p.value
ttc = t.test( subset(sim3, Status == 'NP')$clusters, subset(sim3, Status == 'P')$clusters )$p.value

ggplot(sim3[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + geom_jitter(width=0.2) + 
  labs(title=paste("Norm seg values p-value", signif(ttp, 3)), y='sum similarity')
```

So using the raw segmentation values (not the normalized ones) try some leave one out xval to see if the currently weak differences still hold.

Patients first

```{r xval_pt, echo=T, warning=F, message=T, fig.align='left', fig.height=4, eval=T}
# Leave out patients
loo_xval = matrix(ncol=3, nrow=0, dimnames=list(c(), c('net.similarity', 'sum.similarity', 'clusters')))
for (i in 1:nrow(sum.patient.data)) {
  simL = sum.patient.data[-i,c(2:4)]
  rownames(simL) = sum.patient.data[-i,1]

  for (pt in sum.patient.data$Patient[-i]) {
    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
    ac = apcluster(negDistMat(x1, r=2), q=q)

    simL[pt, 'ns'] = net.similarity(ac)
    simL[pt, 'ss'] = sum.similarity(ac)
    simL[pt, 'clusters'] = length(clusters(ac))
  }

  ttps = t.test( subset(simL, Status == 'NP')$ss, subset(simL, Status == 'P')$ss )$p.value
  ttpn = t.test( subset(simL, Status == 'NP')$ns, subset(simL, Status == 'P')$ns )$p.value
  ttc = t.test( subset(simL, Status == 'NP')$clusters, subset(simL, Status == 'P')$clusters )$p.value

  loo_xval = rbind(loo_xval, cbind(ttpn, ttps, ttc))
}

ggplot(melt(data.frame(loo_xval[,c(1,3)])), aes(x=variable, y=value, fill=variable)) +  
    geom_boxplot() + geom_jitter(shape=16, width=0.2) + geom_hline(yintercept=0.05, color='red', alpha=0.5) +
    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (patient) out clustering')
```

Then randomly leave out samples

```{r xval_smp, echo=F, warning=F, message=T, fig.align='left', fig.height=6, eval=T}

# Leave out samples?
looS_xval = matrix(ncol=3, nrow=0, dimnames=list(c(), c('net.similarity', 'sum.similarity', 'clusters')))
for (i in 1:20) { 
  simX = sum.patient.data[,c(2:4)]
  rownames(simX) = sum.patient.data[,1]

  for (pt in sum.patient.data$Patient) {
      x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-sample(1:(length(patient.data[[pt]]$info$Samplename)), 1)])
      ac = apcluster(negDistMat(x1, r=2), q=q)

      simX[pt, 'ns'] = net.similarity(ac)
      simX[pt, 'ss'] = sum.similarity(ac)
      simX[pt, 'clusters'] = length(clusters(ac))
    }

  ttpn = t.test( subset(simX, Status == 'NP')$ns, subset(simX, Status == 'P')$ns )$p.value
  ttps = t.test( subset(simX, Status == 'NP')$ss, subset(simX, Status == 'P')$ss )$p.value
  ttc = t.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value

  looS_xval = rbind(looS_xval, cbind(ttpn,ttps, ttc))
}

ggplot(melt(data.frame(looS_xval[,c(2,3)])), aes(x=variable, y=value, fill=variable)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
  geom_hline(yintercept=0.05, color='red', alpha=0.5) +
    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (sample) out clustering')
```

## (xval) Re run clusters on the Progressors without the final sample
There is still a difference between the progressors and non, though the p-value does increase. Some of the progressors do start looking more like NPs in their similarity measure but not all of them.

```{r nofinal, echo=F, message=F, warning=F, fig.height=6, fig.width=10}
simX = sum.patient.data[,c(2:4)]
rownames(simX) = sum.patient.data[,1]
# Leave out final sample on the progressors

for (pt in sum.patient.data$Patient) {
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  if (subset(sum.patient.data, Patient == pt)$Status == 'P' & length(which(grepl('HGD', patient.data[[pt]]$info$Pathology))) > 0) {
    final = grep('HGD', patient.data[[pt]]$info$Pathology)
    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-final])
  }
  ac = apcluster(negDistMat(x1, r=2), q=q)

  simX[pt, 'ns'] = net.similarity(ac)
  simX[pt, 'ss'] = sum.similarity(ac)
  simX[pt, 'clusters'] = length(clusters(ac))
}

ttps = t.test( subset(simX, Status == 'NP')$ss, subset(simX, Status == 'P')$ss )$p.value
gg1 = ggplot(melt(data.frame(simX[c('Status','ss')])), aes(x=Status, y=value, fill=Status)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='sim values', x='', title=paste('Leave out HGD sample(s) p-value', signif(ttps, 3)))

ttpc = t.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value
gg2 = ggplot(melt(data.frame(simX[c('Status','clusters')])), aes(x=Status, y=value, fill=Status)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='num. clusters', x='', title=paste('Leave out HGD sample(s) p-value', signif(ttpc, 3)))

grid.arrange(gg1, gg2, ncol=2)

#pander(subset(simX, Status == 'P' & ss <= mean(ss)), caption="These patients are still strongly in the Progressor group")

pander(subset(simX, Status == 'P' & ss >= mean(ss)), caption="These patients are more similar to Non-Progressors without their HGD sample(s)")
```

## Can we identify regions in the progressors?

Not really just from the SVDs.  This isn't any different at the moment from simply calling the variable regions directly.
```{r loc_pca, echo=F, warning=F, message=T, fig.height=12, fig.width=20}
count.occurrences<-function(grs) {
  ov = findOverlaps(grs, type="within", ignoreSelf=TRUE,select="arbitrary")
  counts = countOverlaps(grs, type='within', maxgap=0)

  newgrs = grs[is.na(ov)]

  newgrs$overlaps = 0
  newgrs$patients = ''

  patients = c()  
  index = 1
  occurence = vector(mode='numeric')
  for (x in which(is.na(ov))) {
    occurence = c(occurence, sum(count[index:x]))
    patients = c(patients, paste(grs[index:x]$patient, collapse=','))
    index = x+1
  }
  newgrs$overlaps = occurence
  newgrs$patients = patients

  newgrs$patient.count = unlist(lapply(strsplit(patients, ','), length))

  newgrs$patient = NULL  
  return(newgrs)
}
annotate.cosmic<-function(grs) {
  ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
  ensembl = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
  mart = useDataset("hsapiens_gene_ensembl", ensembl)

  grs$Genes = 0
  grs$Cosmic.count = 0
  grs$Cosmic.genes = ''

  plocs = as.data.frame(grs)  
  for (i in 1:nrow(plocs)) {
    bm = getBM(mart = mart,
        attributes=c('hgnc_symbol', 'entrezgene' ), filters = c('chromosome_name', 'start', 'end' ),
        values = with(plocs[i,], list(gsub("chr", "", as.character(seqnames)), start, end)))

    grs[i]$Genes = length(unique(bm$hgnc_symbol))
    grs[i]$Cosmic.count = length(intersect(bm$entrezgene, ccgenes$Entrez.GeneId))
    grs[i]$Cosmic.genes = paste(intersect(bm$hgnc_symbol, ccgenes$Gene.Symbol), collapse=',')
  }

  return(grs)
}

#pt = 'PR1_HIN_042'
gg.svd<-function(mat, names, sd=0, n=3) {
  df = as.data.frame(mat[,(1:n)])

  for (i in 1:ncol(df)) {
    high = mat[,i] > (mean(mat[,i]) + sd * sd(mat[,i]))
    low = mat[,i] < (mean(mat[,i]) - sd * sd(mat[,i]))

    label = high | low
    color=label

    for (j in 1:length(label))
      label[j] = ifelse(label[j], names[j], '')

    df = cbind(df, label, color)
    colnames(df)[grep('^label$', colnames(df))] = paste('SV',i,'.label',sep='')
    colnames(df)[grep('^color$', colnames(df))] = paste('SV',i,'.color',sep='')
  }
  df$x = 1:nrow(df)
  return(df)
  }

pt.regions = list()
for (pt in sum.patient.data$Patient) {
  print(pt)
  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, Pathology)

  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )

  #plot.pca(pt)
  #pca = prcomp(t(x1), center=T)

  #plot(pca$sdev^2, type='o')

  svd = svd(t(x1))
  colnames(svd$u) = paste('SV',1:ncol(svd$u), sep='') 
  colnames(svd$v) = paste('SV',1:ncol(svd$v), sep='')

  #plot(svd$d, type='o', main='Scree')
  # u matrix
  samples = gg.svd(svd$u, colnames(x1), 0, 3)
  grid.arrange(
    ggplot(samples, aes(y=SV1, x=x, col=SV1.color)) + geom_point() +
      geom_text(aes(label=SV1.label), vjust=0) + guides(color='none'),
    ggplot(samples, aes(y=SV2, x=x, col=SV2.color)) + geom_point() +
      geom_text(aes(label=SV2.label), vjust=0) + guides(color='none'),
    ggplot(samples, aes(y=SV3, x=x, col=SV3.color)) + geom_point() +
      geom_text(aes(label=SV3.label), vjust=0) + guides(color='none'),
    top='SVD Samples')

  # v matrix
  regions = gg.svd(svd$v, rownames(x1), 2, 3)
  grid.arrange(
    ggplot(regions, aes(y=SV1, x=x, col=SV1.color)) + geom_point() +
      geom_text(aes(label=SV1.label), vjust=0) + guides(color='none'),
    ggplot(regions, aes(y=SV2, x=x, col=SV2.color)) + geom_point() +
      geom_text(aes(label=SV2.label), vjust=0) + guides(color='none'),
    ggplot(regions, aes(y=SV3, x=x, col=SV3.color)) + geom_point() +
      geom_text(aes(label=SV3.label), vjust=0) + guides(color='none'), 
    top='SVD Locations')

  locs = as.character(unlist(c(regions[grep('SV1.label',colnames(regions))])))
  locs = unique(locs[grep("\\w+", locs)])

  high = names(which(apply(x1[locs,], 1, function(x) length(which(x >= 1.1))) > 0))
  low = names(which(apply(x1[locs,], 1, function(x) length(which(x <= 0.9))) > 0))

  hls = as.data.frame(do.call(rbind, strsplit(c(high, low), '[p|q]\\.|-')))
  colnames(hls) = c('chr','start','stop')
  hls[c('start','stop')] = lapply(hls[c('start','stop')], function(x) as.numeric(as.character(x)))

  #ac = apcluster(negDistMat(x1, r=2), q=0.2)

  #lapply(ac@clusters, function(x) names(x)[which(names(x) %in% high)])
  #lapply(ac@clusters, function(x) names(x)[which(names(x) %in% low)])

  pt.regions[[pt]] = data.frame(hls, cn=c(rep('gain', length(high)), rep('loss',length(low))))
}
pt.regions = do.call(rbind, pt.regions)

pt.regions$patient = sub('\\..*', '', rownames(pt.regions) )
rownames(pt.regions) = 1:nrow(pt.regions)

prog.grs = sort(makeGRangesFromDataFrame(subset(pt.regions, patient %in% subset(sum.patient.data, Status == 'P')$Patient), keep.extra.columns = T, seqnames.field = 'chr', seqinfo=c(1:22,'X','Y')))

np.grs = sort(makeGRangesFromDataFrame(subset(pt.regions, patient %in% subset(sum.patient.data, Status == 'NP')$Patient), keep.extra.columns = T, seqnames.field = 'chr', seqinfo=c(1:22,'X','Y')))

prog.loss = count.occurrences(subset(prog.grs, grepl('loss',cn)))
prog.gain = count.occurrences(subset(prog.grs, grepl('gain',cn)))

np.loss = count.occurrences(subset(np.grs, grepl('loss',cn)))
np.gain = count.occurrences(subset(np.grs, grepl('gain',cn)))

pg.c = annotate.cosmic(prog.gain)
pl.c = annotate.cosmic(prog.loss)

npg.c = annotate.cosmic(np.gain)
npl.c = annotate.cosmic(np.loss)

with(pg.c,plot(overlaps, Cosmic.count, col='blue', pch=19))
with(pl.c, points(overlaps, Cosmic.count, col='red', pch=19))

with(npg.c, points(overlaps, Cosmic.count, col='green', pch=18))
with(npl.c, points(overlaps, Cosmic.count, col='purple', pch=17))

pander(data.frame('total.p' = sum(length(pg.c), length(pl.c)), 'total.np' = sum(length(npg.c), length(npl.c))), caption="Total regions called as gains/losses for progressors vs non")

t.test( width(pg.c), width(npg.c) ) # Yes for the gains
t.test( width(pl.c), width(npl.c) ) # Not for the losses

@@

### 2017-01-17 Patient Report.Rmd
AP clustering analysis, as far as I’ve gotten so far

  # ORDER segvals by endoscopy year etc
  cur.pt = subset(patient.info, Patient == patient.name)
  segvals[, c(colnames(segvals[1:5]), cur.pt$Samplename)]

>>>>

  # ORDER segvals by endoscopy year etc
  cur.pt = subset(patient.info, Patient == patient.name)
  segvals = segvals[, c(colnames(segvals[1:5]), cur.pt$Samplename)]
